Boundary terms in the action for Causal Sets

### 2014-09-18 test.tex
initial commit

\newcommand{\IGNORE}[1]{}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{aligned}}
\newcommand{\eea}{\end{aligned}}
\newcommand{\mb}[1]{\marginnote{\texttt{\small MB:\,#1}}}
\renewcommand{\ij}[1]{\marginnote{\texttt{\small IJ:\,#1}}}
\newcommand{\ibar}{{\overline{\emph\i}}}
\newcommand{\jbar}{{\overline{\emph\j}}}
\newcommand{\mnote}[1]{\marginpar{\raggedright \scriptsize   #1 } }

\title{Boundary terms in the action for Causal Sets}
 \author[a]{Michel Buck}
 \author[a,b]{\!, Fay Dowker}
 \author[a]{ Ian Jubb\,}
 \author[c]{and Sumati Surya}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
\affiliation[c]{Raman Research Institute, CV Raman Ave, Sadashivanagar, Bangalore 560080, India}
\abstract{ 
We propose a family of formulae for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary. %The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface. 
We show that in the continuum limit one recovers the Gibbons-Hawking-York boundary term in the mean.
}

@@

In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum. This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}. It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(\mathcal{M},g)$ by selecting points in $\mathcal{M}$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold. The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable. This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling. The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions. It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold. If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(\mathcal{M},g)$, then we say that the manifold is a good approximation for the causal set.

@@

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
\be\label{eq:GHYBT_in_continuum}
S_{GHY} = \pm \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\ee
where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$. (We will work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}

Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$. More formally
\be
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$, which corresponds to the rate of change of the surface volume backwards in time, as $n^{\mu}$ is past-pointing. 
\mb{does the antichain analogy really stand?} 
In the causal set, the analogue of a spacelike hypersurface is an anti-chain, i.e. a subset $\mathcal{A}\subseteq\mathcal{C}$ in which all elements are unrelated. The intuitive analogue of the boundary term would then be the rate of change of the number of causal set elements below and above the antichain. We shall see that this intuitive idea indeed bears out. %This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''. \mb{under construction} The analogy These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$.

Consider a causal set $\mathcal C$ obtained by a Poisson sprinkling at density $\rho=1/l^d$ into a $d$-dimensional spacetime $(M,g)$. A Cauchy surface $\Sigma\subset M$ induces a partition $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$ of the sprinkling, where $\mathcal C^+$ and $\mathcal C^-$ denote the restrictions of $\mathcal C$ to the points sprinkled to the future and past of $\Sigma$, respectively. Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of the surface. The most natural definition for the nearest neighbours to the future/past of $\Sigma$ in the sprinkling is to take the minimal/maximal elements in $\mathcal C^\pm$, respectively. Thus $x\in \mathcal C^-$ is a maximal element in $\mathcal C^-$ if it is maximal in the precedence relation of the causal set, i.e. if there exists no element $y\in\mathcal C^-$ such that $x\preceq y$. Figure \ref{fig:Nmin_Nmax} shows an illustration of the idea. The minimal and maximal elements have been highlighted. 
%in COLOUR.

We will also show that it is possible to restrict oneself to either $\mathcal C^+$ or $\mathcal C^-$ and still calculate the GHY term. To do this requires us to define the $k$-next-to-maximal/minimal elements in $\mathcal C^-$ or $\mathcal C^+$ respectively. These are the elements of the causal set that have $k$ elements to their future or past. This means that the $0$-to-maximal/minimal elements are the maximal/minimal elements. 
%Elements with $1$ element to their past/future have been highlighted in COLOUR in Figure \ref{fig:Nmin_Nmax}, and those with $2$ have been highlighted in COLOUR.

Let us denote the number of $k$-next-to-maximal elements in $\mathcal C^-$ by $N_{max}^{(k)}$ and the number of $k$-next-to-minimal elements in $\mathcal C^+$  by $N_{min}^{(k)}$. We propose the following definitions for the discrete Gibbons-Hawking-York boundary term (one can adopt the appropriate definition to best suit their calculation needs):\mb{put $\Sigma$ in argument, $S[\mathcal C, \Sigma]$?}

\be\label{GH_boundary_to_causet}
 \mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\times
  \begin{cases}
   \frac{1}{2}\left(N_{max}^{(0)}-N_{min}^{(0)}\right) \\
   d\: N_{max}^{(1)}-N_{max}^{(0)} \\
   N_{min}^{(0)}-d\: N_{min}^{(1)}
  \end{cases}
\ee
where the constant $c_{d}$ only depends on the spacetime dimension and is given by
\be\label{Cn}
c_{d}=\frac{d(d+1)}{(d+2)}\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{2}{d}},
\ee
$A_d=2\pi^\frac{d+1}2/\Gamma\left(\frac{d+1}{2}\right)$ and denotes the volume of the unit $d$-sphere. In fact, one can write down a general expression for the boundary term as a sum over $k$  $N_{\vphantom{i}max}^{\: (k)}$ and/or $N_{min}^{\: (k)}$ terms with particular coefficients. This can be written as follows:

\begin{gather}\label{general_boundary_sum}
\begin{aligned}
\mathcal{S}^{(d)}_{GHY}[\mathcal C]=\left(l/l_p\right)^{d-2} & c_{d}\left( \sum_m p_m \frac{\Gamma\left(\frac{2}{d}+m \right)}{\Gamma\left(\frac{1}{d}+m \right)}  - \sum_n q_n\frac{\Gamma\left(\frac{2}{d}+n \right)}{\Gamma\left(\frac{1}{d}+n \right)} \right)^{-1} \\
& \times \left( \sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)} N_{max}^{(m)}
+  \sum_n q_n \frac{n!}{\Gamma\left(\frac{1}{d}+n \right)} N_{min}^{(n)}\right)
\end{aligned}
\end{gather}
The coefficients, $p_m , q_n \in \mathbb{R}$, must satisfy the following relation:

\be\label{coefficient_relation}
\sum_m p_m + \sum_n q_n = 0
\ee
The different definitions of the boundary term in (\ref{GH_boundary_to_causet}) are just simple cases of this general formula, and the objects that one would preferably use in a calculation. From above one can see that at least two terms are needed in the sum to find the boundary term. The condition~\eqref{coefficient_relation} means that there will be one free parameter if two terms are present\footnote{For example you could include only the maximal and minimal terms (like the first case in~\eqref{GH_boundary_to_causet}) with $p_0=p$ and $q_0=-p$ where $p\in\mathbb{R}$. This would satisfy~\eqref{coefficient_relation} and would leave $p$ as a free parameter. In this case you would get a factor of $p^{-1}$ from the term after $c_d$ in~\eqref{general_boundary_sum} which will cancel the $p$ from the last term.}. This free parameter will be cancelled by the term after $c_d$ in~\eqref{general_boundary_sum}. Thus, with only two terms, the form of the discrete boundary term is fixed. If one chooses to include more than two terms the coefficients of the $N_{{max\vphantom{i}}}^{\:(k)}$ and/or $N_{{min}}^{\:(k)}$ terms are no longer unique.

To support this proposal we show that in the continuum limit of $l\rightarrow 0$ (i.e. $\rho\rightarrow\infty$) we obtain
%of infinite sprinkling density into a $d$-dimensional spacetime we obtain
\be
\lim_{l\rightarrow0}\langle S^{(d)}_{GHY}[\mathcal C] \rangle= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
\ee
for the most general definition above. Here $\langle\cdot\rangle$ denotes the mean over sprinklings.

We also propose a discrete definition for the spatial volume of the surface:

\be\label{eq:surface_volume}
\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{{\,}_{max/min}}^{\:(0)}
\ee
where the constant $b_d$ is given by

\be\label{constant_b_d}
b_d=d\left[\frac{A_{d-2}}{d(d-1)}\right]^{\frac{1}{d}}
\ee
Again on can find a general expression written as a sum of different terms.

\be\label{general_area_sum}
\mathcal{A}^{(d)}[\mathcal{C}]=\left(l/l_p\right)^{d-1}b_{d}\sum_m p_m \frac{m!}{\Gamma\left(\frac{1}{d}+m \right)}\: N_{{\,}_{max/min}}^{\:(m)}
\ee
where the coefficients, $p_m$, must satisfy $\sum_m p_m =1$. This time we see that only one term is needed to find the surface volume, and that when more than one term is added the coefficients will not be unique. We support this claim in the same way as the boundary term, and show that

\be\label{eq:conjecture_for_area}
\lim_{l\rightarrow0}\langle \mathcal{A}^{(d)}[\mathcal{C}] \rangle= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
\ee
where the right hand side is the spatial volume of the surface $\Sigma$.

%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
\begin{figure}
  \centering
    {\includegraphics[width=\textwidth]{minmaxplot}}
     \caption{An illustration of the idea. Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface. The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
     \label{fig:Nmin_Nmax}
\end{figure}

@@

\subsection{Poisson Sprinklings for $\langle N_{max}^{(m)}\rangle$ and $\langle N_{min}^{(m)}\rangle$}

>>>>

\newcommand{\IGNORE}[1]{}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{aligned}}
\newcommand{\eea}{\end{aligned}}
\newcommand{\mb}[1]{\marginnote{\color{red}{\small MB:\,#1}}}
\renewcommand{\ij}[1]{\marginnote{\color{red}{\small IJ:\,#1}}}
\renewcommand{\ss}[1]{\marginnote{\color{red}{\small SS:\,#1}}}
\newcommand{\ibar}{{\overline{\emph\i}}}
\newcommand{\jbar}{{\overline{\emph\j}}}

\title{Boundary terms in the action for Causal Sets}
 \author[a]{Michel Buck}
 \author[a,b]{\!, Fay Dowker}
 \author[a]{\!, Ian Jubb\,}
 \author[c]{and Sumati Surya}
\affiliation[a]{Theoretical Physics Group, Blackett Laboratory, Imperial College, London, SW7 2AZ, U.K.}
\affiliation[b]{Institute for Quantum Computing, University of Waterloo, ON, N2L 2Y5, Canada}
\affiliation[c]{Raman Research Institute, CV Raman Ave, Sadashivanagar, Bangalore 560080, India}

\abstract{ 
We propose a family of formulae for the boundary term in the action of a causal set that is well-approximated by a continuum manifold with spacelike boundary. %The boundary term is proportional to the difference in the number of elements immediately to future and the number of elements immediately to the past of the surface. 
We show that in the continuum limit one recovers the Gibbons-Hawking-York boundary term in the mean.
}

@@

In order to say we have a causal set analogue for a continuous expression we need a well defined procedure for relating the discrete theory to the continuum. This procedure, or tool, is called a \textit{Poisson sprinkling}, or just a \textit{sprinkling}. It is a Poisson process which provides a way to generate a causet from a $d$-dimensional Lorentzian manifold $(M,g)$ by selecting certain points in $M$ to be the elements of $\mathcal{C}$, with an order relation given by the causal order of the manifold. The number of points chosen in a region of spacetime volume, $V$, is a Poisson random variable. This means that the expected number of points in some region will be $\rho V$, where $\rho$ is the density of the sprinkling. The density is related to the discreteness scale, $l$, by $\rho=l^{-d}$ in $d$ spacetime dimensions. It is called a sprinkling as one can envisage the point selection process as a `sprinkling' of points into the manifold. If a causet, $\mathcal{C}$, can be generated with relatively high probability by a sprinkling into the manifold, $(M,g)$, then we say that the manifold is a good approximation for the causal set.

@@

Consider a sufficiently well-behaved $d$-dimensional spacetime $(M,g)$ and Cauchy surface $\Sigma$ in $M$. The causal past and future sets $M^\pm=J^\pm(\Sigma)$ form a partition of $M$ and $\partial M^\pm = \Sigma$. The Gibbons-Hawking-York boundary term for $M^\pm$ is in this case given by
\be\label{eq:GHYBT_in_continuum}
S_{GHY} = \mp \frac{1}{l_p^{d-2}}\int_{\Sigma} K d\Sigma
\ee
where $l_p$ is the rationalised Planck length and $K$ is the trace of the extrinsic curvature $K_{\mu\nu}=h_{\mu}^\rho h_\nu^\sigma \nabla_\rho n_\sigma$ of $\Sigma$ defined with future-pointing timelike unit normal $n_{\mu}=\partial_\mu S/\sqrt{g^{\mu\nu}\partial_\mu S\partial_\nu S}$. (We will work with a mostly plus convention for the metric, in which this translates into a past-pointing normal vector $n^{\mu}$).\mb{ok or more detail on conventions needed?}

Now observe that the integral in~\eqref{eq:GHYBT_in_continuum} can be thought of as the ``volume gradient'' across the surface $\Sigma$. More formally
\be\label{eq:normal_deriv_boundary}
\int K d\Sigma = \frac{\partial}{\partial n}\int d\Sigma,
\ee
where the right hand side is the derivative of the area $\int d\Sigma$ as each point of $\Sigma$ is moved an equal distance along the unit normal vector, $n^{\mu}$, which corresponds to the rate of change of the surface volume backwards in time, as $n^{\mu}$ is past-pointing. 
\mb{does the antichain analogy really stand?}

We will define the causal set analogue of a spacelike hypersurface to be a particular partition of the set $\mathcal{C}$ into parts $\mathcal{C}^+$ and $\mathcal{C}^-$. For this partition to make sense as a spacelike hypersurface we require that no elements in $\mathcal{C}^+$ can precede any elements in $\mathcal{C}^-$. The reason for this condition is that if one could faithfully embed $\mathcal{C}$ in some manifold, $M$ ($\mathcal{C}$ can be generated via a sprinkling into $M$), then given a spacelike hypersurface, $\Sigma$, we can partition the causet into two bits, denoting the future(past) part as $\mathcal{C}^+$($\mathcal{C}^-$), which is made up of those elements sprinkled into $M^+$($M^-$). The causal relations between the two parts immediately satisfy our requirement for the partition to be spacelike. We will refer to a partition as spacelike if it abides by this condition, even if the causet is not embeddable in some manifold. Spacetime volume in the causal set is obtained simply by counting the number of elements, and hence the volume gradient intuitively corresponds to the difference in the number of ``immediate neighbours'' to the future and to the past of the spacelike partition. What does it mean to be immediate neighbours to a spacelike partition? For immediate neighbours to the past(future) of the partition we mean elements of $\mathcal{C}^-$($\mathcal{C}^+$) that have no elements to their future(past) in $\mathcal{C}^-$($\mathcal{C}^+$) (these elements being called \textit{maximal}(\textit{minimal})). We shall see that this intuitive idea for the boundary term indeed bears out. However, one does not need to know the maximal and minimal elements to calculate the boundary term. It can be found from any elements that lie reasonably close to the surface (in the time coordinate on $M$) but are not necessarily maximal or minimal. This will be made more precise below.

In the continuum the GHY term can be calculated for a spacelike hypersurface, $\Sigma$, that is either a future or past boundary of $M$. We wish to calculate the GHY term for future or past boundaries of causets, but there are a couple issues we must deal with. Firstly, if we can embed a causet $\mathcal{C}$ into $M$ then the boundary surface of $M$ will not induce a partition of $\mathcal{C}$ as it did previously (every element of $\mathcal{C}$ will lie in the bulk of $M$), thus our previous idea for the boundary term does not work any more. Secondly, what does it mean to say a bounding surface of a causet if it is not embeddable in some manifold? Previously if the causet was not embeddable we could still define the surface as a partition of the causet. We shall see that the first issue is not a problem, and we will show that the GHY term can be found for a future(past) boundary of a causet (that is embeddable in $M$) by taking a difference in the number of certain elements that lie close to the future(past) bounding surface of $M$. We state that calculating the GHY term this way must apply whether or not the causet is embeddable, but this runs us into the second issue above. For what boundary are we calculating our GHY term if the causet is not embeddable? We propose that spacelike future and past boundaries exist for all causets of finite cardinality. By definition these causets are finite in temporal extent, and so if they can be embedded in some manifold then that manifold can always have spacelike future and past boundaries. These future and past boundaries are abstract objects for which quantities such as the GHY term can be calculated for. They only have physical meaning when the causet is embeddable in some manifold with a future(past) spacelike boundary, denoted $\Sigma$, in which case the quantities we calculate for the abstract future(past) boundary will agree with those calculated for $\Sigma$ (calculated with the continuum definitions) in some continuum limit.

One can also use partitions of a causet to define the causal set analogues of timelike and null hypersurfaces. If the partition is such that an element of $\mathcal{C}^-$ precedes and element of $\mathcal{C}^+$ which again precedes and element of $\mathcal{C}^-$, then the partition corresponds to a timelike hypersurface. If we have a partition that satisfies the requirements for a spacelike partition, we can define a null partition as that which has the following restrictions. For every $x_0\in\mathcal{C}^+$ one can find a sequence of elements $x_0 \succeq x_1 \succeq ...\succeq x_n$ such that $x_n$ is preceded by nothing and all the $x_i \in \mathcal{C}^+$; and for every $x_0\in\mathcal{C}^-$ one can find a sequence of elements $x_0 \preceq x_1 \preceq ...\preceq x_n$ such that $x_n$ precedes nothing and all the $x_i \in \mathcal{C}^-$. There is also the possibility of a surface changing between spacelike, timelike and null at different parts. In this case the situation is a little more complicated in causal set terms, and one must restrict to a suitable subsets of points such that one of the definitions for the different types of surfaces holds true while the others fail. This subset of points, with its particular relations across the particular partition, then corresponds to a portion of the surface being either spacelike, timelike or null.

%This suggests a very simple analogy for a causal set with a ``spacelike hypersurface''. \mb{under construction} The analogy These unrelated elements are maximal in the sense that there exist no elements $y\in\mathcal{C}\setminus\mathcal{A}$ such that $x\preceq y$, $\forall x\in\mathcal{A}$.

Let us define the operators $N_{\vphantom{i}max}^{(k)}\left[\mathcal{C} \right]$ and $N_{min}^{(k)}\left[\mathcal{C}^- \right]$ that, for a given causet, $\mathcal{C}$, return the number of $k$-next-to-maximal/minimal elements respectively. A $k$-next-to-maximal(minimal) element is one for which there are $k$ elements to its future(past). This means that the $0$-to-maximal(minimal) elements are the maximal(minimal) elements.

Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we propose the following family of causal set operators for calculating the GHY boundary term of this partition:
\be\label{general_boundary_sum}
\mathcal{S}^{(d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p}, \mathbf{q} \right]= \left(l/l_p\right)^{d-2} c_{d}
\left( \sum_m p_m N_{max}^{(m)}\left[\mathcal{C}^- \right]
+  \sum_n q_n N_{min}^{(n)}\left[\mathcal{C}^+ \right]\right)
\ee
where the constant $c_{d}$ only depends on the spacetime dimension and is given by
\be\label{Cn}
c_{d}=\frac{d(d+1)}{(d+2)}\left[\frac{V_{d-1}}{d}\right]^{\frac{2}{d}},
\ee
and where $V_d=\pi^{\frac{d}{2}}/\Gamma\left(\frac{d}{2}+1\right)$ denotes the volume of the unit $d$-ball ($\Gamma(t)$ is a gamma function, defined as $\Gamma(t):=\int_0^\infty dx\: x^{t-1}\: e^{-x}$). We have used $\mathbf{p}$ and $\mathbf{q}$ to stand for infinite dimensional vectors $(p_0,...,p_m,..)$ and $(q_0,...,q_n,..)$ respectively, where all the $p_m,\: q_n \in \mathbb{R}$. These vectors contain a finite number of non-zero entries, corresponding to the coefficients of the different $N_{{max\vphantom{i}}}^{\:(m)}$ and $N_{{min}}^{\:(n)}$ terms. Their values can be chosen arbitrarily, so long as they satisfy the following conditions.
\begin{align}\label{coefficient_relation1}
& \sum_m p_m \frac{\Gamma\left(\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left(\frac{1}{d}+n \right)}{n!}=0
\\
& \label{coefficient_relation2}\sum_m p_m \frac{\Gamma\left(\frac{2}{d}+m \right)}{m!}  - \sum_n q_n\frac{\Gamma\left(\frac{2}{d}+n \right)}{n!}=1
\end{align}
The aforementioned family simply corresponds to all the different vectors, $\mathbf{p}$ and $\mathbf{q}$, that satisfy the above criteria.

Given a $d$ dimensional, spatially compact Lorentzian manifold, $M$, the family of operators $\mathcal{S}^{(d)}_{GHY}$ (for different $\mathbf{p}$ and $\mathbf{q}$) defines a family of random variables in the following way. The process of sprinkling points into $M$ with density $\rho=l^{-d}$ generates a random causet $\mathcal{C}$, and so the operators $N_{\vphantom{i}max}^{(k)}$ and $N_{min}^{(k)}$ acting on this random causet are now the random variables $\textbf{N}_{\vphantom{i}max}^{(k)}$ and $\textbf{N}_{min}^{(k)}$. These random variables can be substituted into~\eqref{general_boundary_sum} for $N_{\vphantom{i}max}^{(k)}$ and $N_{min}^{(k)}$ to give the family of random variables, $\textbf{S}^{(d)}_{GHY}$. These random variables will be functions of $\rho$ and the two parts of the manifold, $M^{\pm}$, where  $\textbf{N}_{\vphantom{i}max}^{(k)}$($\textbf{N}_{min}^{(k)}$) depends on $M^-$($M^+$) and $\rho$. The expectation value of $\textbf{S}^{(d)}_{GHY}$ in this sprinkling process will be denoted by $\overline{S}^{(d)}_{GHY}:=\mathbb{E}(\textbf{S}^{(d)}_{GHY})$. We claim that in the limit of infinite density the value of $\overline{S}^{(d)}_{GHY}$, for any vectors $\mathbf{p}$ and $\mathbf{q}$ satisfying~\eqref{coefficient_relation1} and ~\eqref{coefficient_relation1}, tends to the continuum GHY boundary term for a surface $\Sigma$ that partitions the manifold into $M^+$ and $M^-$:
\be
\lim_{l\rightarrow0}\overline{S}^{(d)}_{GHY}[M^-,M^+,\rho;\mathbf{p} , \mathbf{q}]= \frac{1}{l_p^{d-2}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}\: K\label{eq:mainconjecture}
\ee

There is a huge amount of freedom that remains in the choice of $\mathbf{p}$ and $\mathbf{q}$. One can see that at least two terms are necessary in order to satisfy~\eqref{coefficient_relation1} and~\eqref{coefficient_relation2}. For any two terms the coefficients will be unique, and for more than two terms this uniqueness is lost. In a finite difference method one can form a discrete derivative in one direction by taking a difference of two values at different points along that direction, and dividing by the distance between them. We are doing something very similar here. The boundary action (in the continuum) is a derivative in one direction, the normal direction (as in~\eqref{eq:normal_deriv_boundary}), and so a finite difference method should be able to approximate this with just two terms. This agrees with the fact that we are able write the discrete boundary action with only two terms.

There are a few, two-term, members of the family that are worth mentioning, as they have simple forms that make them good candidates for computations. Given a causet, $\mathcal{C}$, with a spacelike partition so that $\mathcal C = \mathcal C^+ \cup\, \mathcal C^-$, we define the causal set operator $\mathcal{S}^{(d)}_{\pm}$ as a particular member of the family in~\eqref{general_boundary_sum}:
\be\label{eq:min_max_simple_formula}
\mathcal{S}^{(d)}_{\pm}\left[\mathcal{C}^-,\mathcal{C}^+\right]:=\mathcal{S}^{(d)}_{GHY}\left[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p}, -\mathbf{p} \right]=\left(l/l_p\right)^{d-2}\frac{c_{d}}{2\Gamma\left(\frac{2}{d} \right)}\left( N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C}^- \right] - N_{min}^{\: (0)}\left[\mathcal{C}^+ \right] \right)
\ee
where $\mathbf{p}=\left(\left(2\Gamma\left(\frac{2}{d} \right)\right)^{-1},0,0,...\right)$.
This expression is the easiest to employ computationally, and we shall use its random variable counterpart, $\textbf{S}^{(d)}_{\pm}\left[M^-,M^+,\rho\right]:=\textbf{S}^{(d)}_{GHY}\left[M^-,M^+,\rho;\mathbf{p}, -\mathbf{p} \right]$, later when we study the fluctuations of the discrete boundary actions numerically.

If one wants to calculate the GHY boundary term for the future(past) boundary of a single causet, $\mathcal{C}$, then one simply inserts $\mathcal{C}$ into the $\mathcal{C}^-$($\mathcal{C}^+$) argument of $\mathcal{S}^{(d)}_{GHY}$ and sets $\mathbf{q}$($\mathbf{p}$) to the zero vector, $\mathbf{0}=(0,0,...)$. This renders the $\mathcal{C}^+$($\mathcal{C}^-$) argument unnecessary and so one can insert the empty set, $\emptyset$, for convenience. We define the operators for the future and past boundaries respectively as:
\begin{align}\label{eq:future_past_boundary_terms1}
\mathcal{S}^{(d)}_{+}[\mathcal{C}]:= & \mathcal{S}^{(d)}_{GHY}\left[\mathcal{C},\emptyset;\mathbf{p},\mathbf{0}\right] =\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\left( d\:N_{\vphantom{i}max}^{\: (1)}\left[\mathcal{C} \right] - N_{\vphantom{i}max}^{\: (0)}\left[\mathcal{C} \right] \right)
\\
\label{eq:future_past_boundary_terms2}
\mathcal{S}^{(d)}_{-}[\mathcal{C}]:= & \mathcal{S}^{(d)}_{GHY}\left[\emptyset,\mathcal{C};\mathbf{0},-\mathbf{p}\right] =\left(l/l_p\right)^{d-2}\frac{c_{d}}{\Gamma\left(\frac{2}{d} \right)}\left( N_{min}^{\: (0)}\left[\mathcal{C} \right] - d\: N_{min}^{\: (1)}\left[\mathcal{C} \right] \right)
\end{align}
where $\mathbf{p}=\left(-\Gamma\left(\frac{2}{d} \right)^{-1},d\:\Gamma\left(\frac{2}{d} \right)^{-1},0,0,...\right)$. In writing down an action for a particular causal set, $\mathcal{C}$, one would write down the Benincasa-Dowker action, $\mathcal{S}^{(d)}_{BD}[\mathcal{C}]$. To complete the story we propose that one must include the boundary terms $\mathcal{S}^{(d)}_{+}[\mathcal{C}]$ and $\mathcal{S}^{(d)}_{-}[\mathcal{C}]$. Explicitly, the total action for a causal set should be given by:
\be\label{eq:total_causet_action}
\mathcal{S}^{(d)}=\mathcal{S}^{(d)}_{BD}+\mathcal{S}^{(d)}_{+}-\mathcal{S}^{(d)}_{-}
\ee
where we have omitted the causet argument, $\mathcal{C}$.

There may be contributions from the causal set analogue of timelike boundaries (although there are arguments to suggest that they make no contribution), and we discuss below certain attempts we have made to find discrete GHY terms for such boundaries.

We also propose a family of operators for the causal set analogue of spatial volume for the spacelike partition:
\be\label{general_area_sum}
\mathcal{A}^{(d)}[\mathcal{C}^-,\mathcal{C}^+;\mathbf{p},\mathbf{q}]=\left(l/l_p\right)^{d-1}b_{d}\left(\sum_m p_m N_{\vphantom{i}max}^{\: (m)}\left[\mathcal{C}^- \right] + \sum_n q_n N_{min}^{\: (n)}\left[\mathcal{C}^+ \right]\right)
\ee
where the constant $b_d$ is given by
\be\label{constant_b_d}
b_d=d\left[\frac{V_{d-1}}{d}\right]^{\frac{1}{d}}
\ee
and we have used similar definitions for the arguments $\mathbf{p}$ and $\mathbf{q}$. This time however, the coefficients must satisfy 
\be\label{area_coefficient_relation}
\sum_m p_m \frac{\Gamma\left(\frac{1}{d}+m \right)}{m!}  + \sum_n q_n\frac{\Gamma\left(\frac{1}{d}+n \right)}{n!}=1
\ee
From this relation we can see that only one term is necessary to determine an expression for the discrete surface volume, and that when more than one term is added the coefficients will cease to be unique. Once again we can define a family of random variables, $\textbf{A}^{(d)}$, and their expectation values, $\overline{A}^{(d)}:=\mathbb{E}(\textbf{A}^{(d)})$, which will depend upon $M^{\pm}$ and $\rho$. We make another claim that, in the limit of infinite density, $\overline{A}^{(d)}$ tends to the spatial volume of the surface $\Sigma$.
\be\label{eq:conjecture_for_area}
\lim_{l\rightarrow0}\overline{A}^{(d)}[M^-,M^+,\rho;\mathbf{p} , \mathbf{q}]= \frac{1}{l_p^{d-1}}\int_{\Sigma} d^{d-1}x\: \sqrt{h}
\ee
where the right hand side is the spatial volume of the surface $\Sigma$. In proving~\eqref{eq:mainconjecture} we will establish the results necessary to prove this claim as well.

By making the same restrictions on the arguments as we did for $\mathcal{S}^{(d)}_{GHY}$, we can find expressions for the causal set analogue of the spatial volume of the future or past boundaries. We define operators for the future and past boundaries respectively as two of the simplest members of the family:
\begin{align}\label{eq:future_past_spatial_volume}
\mathcal{A}^{(d)}_{+}[\mathcal{C}]:=\mathcal{A}^{(d)}[\mathcal{C},\emptyset;\mathbf{p},\mathbf{0}]= & \left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{\vphantom{i}max}^{\: (0)}[\mathcal{C}]
\\
\mathcal{A}^{(d)}_{-}[\mathcal{C}]:=\mathcal{A}^{(d)}[\emptyset,\mathcal{C};\mathbf{0},\mathbf{p}]= & \left(l/l_p\right)^{d-1}\frac{b_{d}}{\Gamma\left(\frac{1}{d}\right)}\: N_{min}^{\: (0)}[\mathcal{C}]
\end{align}
where $\mathbf{p}=(\Gamma\left(\frac{1}{d}\right)^{-1},0,0,...)$.

%MB1: (too vague) During the calculation approximations will be made. These approximations are justified on the grounds that if certain orders are retained throughout the calculation they will in fact vanish in the limit $\rho \rightarrow \infty$
%\begin{figure}
  %\centering
   % {\includegraphics[width=\textwidth]{minmaxplot}}
    % \caption{An illustration of the idea. Pictured here is a causal set obtained by a sprinkling into a spacetime partitioned by a spacelike hypersurface. The minimal and maximal points about the surface are highlighted and the shaded regions illustrate the regions whose volumes $V_\blacktriangle$ and $V_\blacktriangledown$ are needed in the proof.}
     %\label{fig:Nmin_Nmax}
%*\end{figure}

@@

\subsection{Poisson Sprinklings for $\left\langle N_{max}^{\:(m)}\right\rangle$ and $\left\langle N_{min}^{\: (n)}\right\rangle$}

In order to prove~\eqref{eq:mainconjecture} and~\eqref{eq:conjecture_for_area} let us first derive expressions for the mean values of $N_{\vphantom{i}max}^{\: (m)}$ and $N_{min}^{(n)}$. The proof will will follow almost automatically, as~\eqref{general_boundary_sum} and~\eqref{general_area_sum} are just linear combination of $N_{\vphantom{i}max}^{\: (m)}$ and $N_{min}^{(n)}$ terms, and thus their mean values will just be linear combinations of $\left\langle N_{\vphantom{i}max}^{\: (m)}\right\rangle$ and $\left\langle N_{min}^{\: (n)}\right\rangle$ terms.
